---
title: "Numeric EDA"
subtitle: "Part II"
output: github_document
date: "2022-10-04"
---

```{r setup}
library(tidyverse)
```

## Let's import data

```{r}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

```

Floor_date where unit = month -->  rounds to the bottom of the month


## `group_by()`

```{r}
weather_df %>% 
  group_by(name, month)
```


## `summarize()`

let's group and then summarize!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n()
  )
```


can we count in other ways?

```{r}
weather_df %>% 
  count(name, month)
```


don't use `table()` -- it's not a dataframe. If you like this, don't. You can't do anything with it.

```{r, eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```


you can have multiple summaries

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_dist = n_distinct(date)
  )
```


## a brief aside on 2x2 tables

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  pivot_wider(
    names_from = cold,
    values_from = n_obs
  )
```


quicker and easier way to generate a 2x2 table: 

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```



## General summaries

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    sd_prcp = sd(prcp),
    median_tmax = median(tmax, na.rm = TRUE)
  )
```


Summary in more compact way -- do it `across` variables
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```


This is a dataframe!!
You can use this to plot...

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_path()
```


```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digit = 2)
```

Renders a formatted table for markdown document



