---
title: "Visualization"
subtitle: "Part I"
output: github_document
date: "2022-09-29"
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data
Creating a data frame using the `meteo_pull_monitors` in `rnoaa` (allows you to access public data online that is aggregated); using this to pull data rather than creating it ourselves or reading a CSV
```{r}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots!!

Create my first scatterplot ever.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

New approach, same plot: Piping

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Can also add mutate, filter, or select --> can do data manipulation before you make the plot, but all in a row


Save and edit a plot object.

```{r}
weather_plot <-
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot
#can treat this like an existing plot that you can add to.

weather_plot + geom_point()
```

## Advanced scatterplot...

#### Start with the same one and make it fancy

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

Smooth curves that run through the middle of the dataset. Helpful to know that `geom_smooth` is additional geometry you can add when you're exploring your data.


#### What about the `aes` placement..?
Can specify some aesthetics within geom_point. It ends up looking the same

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```
Instead of having one line for each `name`, we just have a sinngle blue line that runs through the data. Reason for that is because color is only applied to the scatterplot when it was defined within the `geom_point` argument. 

Be aware that you can define aesthetics in ggplot, which makes it global and will be passed through everything else. 


#### Let's facet some things!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

First argument defines rows ~ second argument defines columns.
Creates three paneled plot with smooth curves. 


#### Adjusting the alpha level and the transparency:

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .2) +
  geom_smooth(se = FALSE, size = 1) + 
  facet_grid(. ~ name)
```
Defining aesthetics related to specific variables

* alpha = .5 > 50% transparency
  * If you define alpha outside of the aesthetic, meaning not applying it to a variable, you can define it globally by including it in another function (using geom_point like shown here rather than using ggplot function alpha = tmin or something)

#### Let's combine some elements and try a new plot. 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```



## Some small notes

How many geoms have to exist?

You can have whatever geoms you want. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
# Hexagons
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()

# Squares
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_bin2d()

# Like an elevation map
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_density2d() +
  geom_point(alpha = 0.3)

```

Helpful when you have a plot that shows a LOT of data points and you want to emphasize certain things


## Univariate plots

Histograms are really great.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color ...

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram()

# Solves position argument (not overlapping)
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram(position = "dodge")

# Filling bars
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")

# Could do one at a time by faceting
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)

# Removing fill argument; don't necessarily need to color whenn faceting
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Use a different geometry if you don't like it / want to compare distributions across levels..

So, let's try a new geometry!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.4, adjust = 0.5)
```

Density = histogram that's been smoothed out around the edges. You do lose a little bit of information. You can make some info come back if you play with the bin width (adjust = 0.5). 

This is just an exploration so no need to be perfect. 

What about boxplots?

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```


Trendy plots :-)

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5)
```

But this isn't showing me the median.

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5) +
  stat_summary(fun = "median")
```

Ridge plots... the most popular plot of 2017

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```


