---
title: "Visualization"
subtitle: "Part II"
output: github_document
date: "2022-10-04"
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data

```{r}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

## Scatterplot

But better this time

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5)
```

We want to tidy this up and add some features:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  )
```

Playing around with various scales -- ggplot will make choices about the mapping of the aesthetics, but that can be changed.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10C", "0", "15")
  ) + 
  scale_y_continuous(
    trans = "sqrt"
  )
```

NOTE: Sqrt and log transformations can be helfpul to be able to do since they come up

Make a new version: 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100, 300))
```

Instead of using hue, use this:

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```

Viridis works great for people even if they have color blindness
