---
title: "Visualization"
subtitle: "Part II"
output: github_document
date: "2022-10-04"
---

```{r setup}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## Load the weather data

```{r}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

## Scatterplot

But better this time

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5)
```

We want to tidy this up and add some features:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  )
```

Playing around with various scales -- ggplot will make choices about the mapping of the aesthetics, but that can be changed.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10C", "0", "15")
  ) + 
  scale_y_continuous(
    trans = "sqrt"
  )
```

NOTE: Sqrt and log transformations can be helfpul to be able to do since they come up

Make a new version: 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100, 300))
```

Instead of using hue, use this:

```{r}
ggplot_weather <- weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```

Viridis works great for people even if they have color blindness


## Themes

```{r}
ggplot_weather +
  theme_minimal() +
  theme(legend.position = "bottom")
```

Also, theme_bw, theme_classic, etc. 
Recognize that order of the theme commands matter -- putting legend.position before theme_minimal removes the position of the legend that you just fixed


## Setting Options
First code chunk of any markdown document has extra lines that details what the plots should look like. This is an example (what Jeff does):

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



## Data in `geom()`

```{r}
central_park_df <- 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki_df <- 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(waikiki_df, aes(x = date, y = tmax)) +
  geom_point() +
  geom_line(data = central_park_df)

```

## Patchwork...

```{r}
tmax_tmin_plot <-
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  theme(legend.position = "none")

prcp_density_plot <-
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) + 
  theme(legend.position = "none")

tmax_tmin_plot
prcp_density_plot

tmax_tmin_plot + prcp_density_plot
tmax_tmin_plot / prcp_density_plot

seasonality_plot <-
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

(tmax_tmin_plot + prcp_density_plot) / seasonality_plot
```


## Data manipulation

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

If you're making a plot, whatever is on the X-axis is converted to a factor (since it has to be a character variable) -- then default is that it's made alphabetical. 

How to change this:

```{r}
weather_df %>% 
  mutate(name = fct_relevel(name, "Waikiki_HA")) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

You can say to start with one and then it goes alphabetical after that


Reorder one variable according to values of a different variable: `fct_reorder`
(most likely using the median to order)

```{r}
weather_df %>% 
  mutate(name = fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Takeaways:

* If you're dealing with categorical predictors -- R is thinking of them as factors
* If you want to change the order -- you need to manipulate the data -- don't fight ggplot to make it happen
* Factors matter a lot; we'll worry more about them later on


## Example / Importance of Tidying/Data Manipulation 

```{r}
pulse_df <-
  haven::read_sas('data/public_pulse_data.sas7bdat') %>% 
  janitor::clean_names()
```
This dataset is not structured in the way you need to make a boxplot; you need 'visit' as a variable in the dataset.

PROBLEM IN MAKING THAT FIGURE: the data isn't structured how it needs to be to make the figure that you want it to be.


```{r}
pulse_df <-
  haven::read_sas('data/public_pulse_data.sas7bdat') %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  ) %>% 
  select(id, visit, everything()) %>% 
  mutate(
    visit = fct_relevel(visit, "bl")
  )

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()

```

